# CTLSC0x - Supply Chain Analytics

```{r include=FALSE}
library(ompr)
library(ROI)
library(ROI.plugin.glpk)
library(ompr.roi)
library(ggplot2)
library(plotly)
library(tibble)
library(DT)
library(tidyverse)
library(knitr)
```


## Week2 - Unconstrained & Constrained Optimization

### Recitation

#### Unconstrained

Straigthforward exercises in modeling optimization without index vectors.

$$
\textbf{Equation 1:} \ \ Minimze \ \ y = x^2+2*x-3
$$

Since is an unconstrained quadratic optimization with a single variable, base R package optimizer can solve it

```{r}
f <- function(x){x^2+2*x-3}
result <- optimize(f,interval = c(-10,10), maximum = F)
print(result)

```

Let's check the result in the plot

```{r}
#Data created for plot, using mutate on the function
x <- seq(-10,10,1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$minimum)

```



$$
\textbf{Equation 2:} \ \ Maximize \ \ z =-x^2+2*x-y^2
$$

In this equation there is 2 variables, for this case it can be used the base R package optim.


```{r}
f <- function(x,y){-x[1]^2+2*x[1]-x[2]^2}
result <- optim(c(1, 1), f)
print(result)

```


#### Constrained

##### Banner Chemicals

```{r}

data <- tibble(Type = c("Profit","Plant","Additive A", "Additive B"), High = c(80,1,3,1),
                   Supreme = c(200,1,2,3), Capacity = c(NA,110, 300, 280))

datatable(data)

```



$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i\in M}^{ }p_i\cdot x_i \\
\textbf{subject to}\\
    & \sum_{i\in M} x_i \le\ C \\
    & \sum_{i\in M} a_{i,j} \cdot x_i \le\ A_j, \ \ \forall j \in N \\
    & x_i \ge 0 \\
    \\
    
\textbf{Where} \\

& i = \text{Products in M} \\
& j = \text{Additives in N} \\
& p_i = \text{Profit margin for product i} \\
& C = \text{Plant capacity} \\
& A_j = \text{Additive j available} \\
& a_{i,j} = \text{Quantity of additive j required per barrel of product i} \\
& x_i = \text{Quantity of product i to produce} \\

\end{align*} 
$$

```{r, warning=FALSE}
rm(x)

n = 2 #Number of products High and Supreme
C = filter(data, Type == "Plant") %>% select(Capacity) %>% .$Capacity
a = as.matrix(filter(data, Type %in% c("Additive A","Additive B")) %>% select(High, Supreme))
A = filter(data, Type %in% c("Additive A","Additive B")) %>% select(Capacity) %>% .$Capacity
p = filter(data, Type == "Profit") %>% select(High, Supreme) %>% gather() %>% .$value


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # minimize travel distance
  set_objective(sum_expr(p[i]*x[i], i = 1:n), "max") %>%

  # you cannot exceed the Plant Capacity
  add_constraint(sum_expr(x[i], i = 1:n) <= C) %>%
  
  #You cannot exceed aditivie capacity
  add_constraint(sum_expr(x[i] * a[j,i], i = 1:n) <= A[j], j = 1:n)

#How to check for the constraints
model$constraints

#How to check for the Objective
model$objective 

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
#get_column_duals(result) Not working

```


### Practice Problems

#### Handmade Baskets

$$
P = (x/73)^2-(x/95)+0.92
$$

```{r}
P <- function(x){(x/73)^2-(x/95)+0.92}

#Data created for plot, using mutate on the function
x <- seq(-100,200,1)
data <- data <- data.frame(x) %>%
        mutate(y = P(x))

min = data[data$y == min(data$y),] #cheating :P

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = min$y, x = min$x)

```




```{r}
#Part 1: Price Functio

#Question1: How much will it charge with 5 basket's order ?
P(5)

#Question2: What is the first derivative of the price function?
library(Ryacas)
x <- Sym("x")
s <- expression((x/73)^2-(x/95)+0.92)
deriv(s,x) #Didn't work TO DO

#Question3: Second Derivative
x <- Sym("x")
s <- expression(-1/95 + (2*x)/5329)
deriv(s,x) #worked

#Question4: Lowest Unit Price
result <- optimize(P,interval = c(-100,100), maximum = F)
print(result)

```

#### Maximizing Storage


$$
\begin{align*}
\textbf{Minimize} & \ Z = x \cdot y \\
\textbf{Subject to} \\
& 2 \cdot x+2 \cdot y = 26.7
\end{align*}
$$


Simplifying $2*x+2*y = 26.7$ as $y = (26.7 - 2*x)/2$ is possible to change the optimization problem to a single variable:

$$
\begin{align*}
\textbf{Minimize} & \ Z\ =x\cdot\ \left(\frac{\left(26.7-2\cdot x\right)}{2}\right)\\
\end{align*}
$$


Solving:

```{r}
f <- function(x){x*((26.7 - 2*x)/2)}
result <- optimize(f,interval = c(-100,100), maximum = T)
print(result)

```


Let's check the result in the plot

```{r}
#Data created for plot, using mutate on the function
x <- seq(-100,100,0.1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)

```


#### Marketing MagicPuppy

$$
N = 1700 \left ( \frac{x}{90} - \left ( \frac{x}{90} \right ) ^2 \right )
$$



```{r}
f <- function(x){1700*(x/90-(x/90)^2)}

#Question1: First derivative
x <- Sym("x")
s <- expression(1700*(x/90-(x/90)^2))
deriv(s,x) #Didn't work TO DO


#Question2: How many units should be given away in the campaign in order to maximize its positive impact?
result <- optimize(f,interval = c(-100,200), maximum = T)
print(result)


x <- seq(-50,100,1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)

```

#### CureBase Cancer Institute


$$
W = \left ( m^{x^{2}} \right )\left ( n^{-x} \right )\left ( z^7 \right )
$$


```{r}
#Question what is the value of  x  at the maximum value of  W , given that  m=0.06 ,  n=0.14 ,  z=0.15 ?)
f <- function(x){0.06^(x^2)*(0.14^(-x))*0.15^7}
result <- optimize(f,interval = c(-1,1), maximum = T)
print(result)
```



```{r}
x <- seq(-1,1,0.1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)
```

 
#### Santa’s Molding Company


```{r}

data <- tibble(Type = c("Profit","Time","Cubic Area", "index capacity"),
               Toy = c(25,13,33,75), doll = c(25,15,25,70),
               Capacity = c(NA,1440, 3000, NA))

datatable(data)

```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i\cdot x_i \\
\textbf{subject to}\\
    & x_i \le\ c_i  \ \ \forall i \in N \\
    & \sum_{i=1}^{N} t_{i} \cdot x_i \le\ 1440 \\
    & \sum_{i=1}^{N} a_{i} \cdot x_i \le\ 3000 \\
    & x_i \ge 0 \\
    \\
    
\textbf{Where} \\

& i = \text{Products in N} \\
& p_i = \text{Profit margin for product i} \\
& c_i = \text{Product daily capacity} \\
& t_i = \text{Time to mold product i} \\
& a_i = \text{Cubic inches of product i} \\
& x_i = \text{Quantity of product i to mold} \\

\end{align*} 
$$

If we look closer on the dataframe, we can set constraints 2 and 3 as one with a matrix including capacity constraints, but for math notation I've set them differently since they are 2 different types of entities.


```{r}
rm(x)

n = 2 #Number of products Toy and doll
c = filter(data, Type == "index capacity") %>% select(Toy,doll) %>% gather() %>% .$value #index_capacity
a = as.matrix(filter(data, Type %in% c("Time","Cubic Area")) %>% select(Toy,doll)) #constraints left hand
C = filter(data, Type %in% c("Time","Cubic Area")) %>% select(Capacity) %>% .$Capacity #constraints right hand
p = filter(data, Type == "Profit") %>% select(Toy,doll) %>% gather() %>% .$value #profit



model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # minimize travel distance
  set_objective(sum_expr(p[i]*x[i], i = 1:n), "max") %>%

  # you cannot exceed the Plant Capacity
  add_constraint(x[i] <= c[i], i= 1:n) %>%

  #You cannot exceed aditivie capacity
  add_constraint(sum_expr(x[i] * a[j,i], i = 1:n) <= C[j], j = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = FALSE))

#Optimal Value
result$objective_value

#result solution
result$solution

```


Let's solve the questions

```{r}
P <- function(x,y){25*x+25*y}
#Question 1: Graphical selection
#Question 2: What is the value of x = 70 and y= 30
P(70,30)

#Question 3: What is the value of x = 50 and y= 53
P(50,53)

#Question 3: What is the value of x = 50 and y= 53
P(20,75)

```

#### Crazy Cereal

```{r}

data <- tibble(Type = c("Sugary","Regular","Capacity"), Sugar = c(0.66,0.21,2000),
                   Corn_Flake = c(0.34,0.79,4000), Profit = c(0.94,0.82,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le C_j, \ j \in M \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    \\
    
\textbf{Where} \\

& i = \text{Products in N} \\
& j = \text{Components of N = M} \\
& p_i = \text{Profit margin for product i} \\
& a_{i,j} = \text{components for product i} \\
& C_j = \text{Component capacity} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake

#Let's do some roots R instead of dplyr :)
C = as.matrix(data[data$Type == "Capacity",c("Sugar", "Corn_Flake")])
a = as.matrix(data[data$Type %in% c("Sugary","Regular"),c("Sugar", "Corn_Flake")])
p = as.matrix(data$Profit[1:2])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(p[i] * x[i], i = 1:n), "max") %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:m) <= C[j], j = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = FALSE))

#Optimal Value
result$objective_value

#result solution
result$solution

```

#### Jim's Meat Packing Company

```{r}

data <- tibble(Cut = c("Chuck","Sirloin","Restriction"), Lean_Meat = c(0.09,0.6,0.3),
                   Fat_Meat = c(0.02,0.06,0.05), Cost = c(9.3,8.4,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=1 \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=2 \\
    & \sum_{i=1}^{N} x_{i} \ge 50 \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
    \\
    
\textbf{Where} \\

& i = \text{Type of cut in N} \\
& j = \text{Type of meat} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& R_j = \text{Restriction capacity} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake

#Another way of subsetting data in R
R = as.matrix(data[3,2:3])
a = as.matrix(data[1:2,2:3])
c = as.matrix(data[1:2,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #yay
  add_constraint(sum_expr(x[i], i = 1:n) >= 50) %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

  
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2)


#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)


```


I couldn't find a proper function on shadow price calculation on ompr package, so lets try to calculate it ourselves!

```{r}

ModelIter <- function(z){
  
  model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #yay
  add_constraint(sum_expr(x[i], i = 1:n) >= z) %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2) %>%
    
  solve_model(with_ROI(solver = "glpk", verbose = F))
  
  return(model$objective_value)
  
}

Iter_Results <- tibble(constraint = seq(48,55,1)) %>%
                mutate(Obj_value = as.numeric(map(constraint, ModelIter)),
                       Marginal_value = round((Obj_value-lag(Obj_value,1)/(constraint-lag(constraint,1))),2))

datatable(Iter_Results)
mean(Iter_Results$Marginal_value[2:length(Iter_Results)]) #Didn't work

```

Marginal Value set as 8.85 got us an error, the answer was 8.63, would be good to understand what was made wrong in here.


#### John's Shipping Company

```{r}

data <- tibble(Fuel_Type = c("O_Type","D_Type","Restriction"), Hydrogen_Conc = c(45,90,60),
                   Oxygen_Conc = c(15,6,9), Cost = c(1.05,1.34,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=1 \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=2 \\
    & \sum_{i=1}^{N} x_{i} \ge G \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
    \\
    
\textbf{Where} \\

& i = \text{Type of cut in N} \\
& j = \text{Type of meat} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& R_j = \text{Restriction capacity} \\
& x_{i} = \text{Quantity of product} \\
& G = \text{Minimum to be produced} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake
z = 10000 #minimum to be produced

#Another way of subsetting data in R
R = as.matrix(data[3,2:3])
a = as.matrix(data[1:2,2:3])
c = as.matrix(data[1:2,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #Produce at least the minimum required
  add_constraint(sum_expr(x[i], i = 1:n) >= z) %>%

  #Minimum component required
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

   #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

#Optimal Value
result$objective_value

#result solution
result$solution


```


#### A New Office Plant

```{r}

data <- tibble(Type = c("Product_1","Product_2","Product_3", "Min", "Max"),
               Nitrogen  = c(360,380,310,1800,2200), Potassium = c(30,20,20,100,100),
               cost = c(1.59,2.19,2.99,NA,NA))

datatable(data)

```



$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge Min_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le Max_j, \ \ j \in M \\
    & x_{i} \ge 0 \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Type of component in M} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& Min_j = \text{Restriction of component} \\
& Max_j = \text{Restriction of component} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
min = as.matrix(data[4,2:3])
max = as.matrix(data[5,2:3])
a = as.matrix(data[1:3,2:3])
c = as.matrix(data[1:3,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "continuous", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #Produce at least the minimum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) >= min[j], j = 1:m) %>%
  
  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j], j = 1:m)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

#Optimal Value
result$objective_value

#result solution
result$solution


#Question: How many mg of Nitrogen would your solution provide to the plant every day? (Please round your answer to integer number)
result$solution[1]*a[1,1]+result$solution[2]*a[2,1]

```


#### Rochak's Ink Company

```{r}

data <- tibble(Type = c("P_Cyan","P_Magenta","P_Yellow","Max"),
               Stage1  = c(0.44,0.51,0.5,17*60^2), Stage2 = c(0.55,0.6,0.59,23*60^2),
               Demand = c(84000,72000,93000,NA), Profit = c(0.6,0.72,0.62,NA))


datatable(data)
```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le Max_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \le D_i \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Production stage in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Time requirement for product i in stage j} \\
& Max_j = \text{Restriction of production time} \\
& D_i = \text{Demand for product} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$

In this exercise, you're willing to maximize profit given the demand. You don't have to fulfill all of the demand, that's why the $ \sum_{i=1}^{N} x_{i} \le D_i$ constraint.



```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
D = as.matrix(data[1:3,4])
max = as.matrix(data[4,2:3])
a = as.matrix(data[1:3,2:3])
p = as.matrix(data[1:3,5])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(p[i] * x[i], i = 1:n), "max") %>%
  
  #Don't produce something you don't have demand for
  add_constraint(x[i] <= D[i], i = 1:n) %>%
  
  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j], j = 1:m)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

#Optimal Value
result$objective_value

#result solution
result$solution

```

#### Pasteur Cheese Factory


```{r}

data <- tibble(Type = c("P_Original","P_Cheddar","P_Wasabi","Max","Cost"),
               X_Ray  = c(8.6,2.2,5.1,8*60,0.11), Checkweigher = c(7.0,6.8,9.8,9*60,0.22),
               Profit = c(20,15,17,NA,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} x_{i} \cdot \left( p_i - \sum_{j=1}^{M} c_{i,j} \right) \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le Max_j, \ \ j \in M \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Production stage in M} \\
& p_i = \text{Profit for product i} \\
& c_{i,j} = \text{Cost for product i in stage j} \\
& a_{i,j} = \text{Time requirement for product i in stage j} \\
& Max_j = \text{Restriction of production time} \\
& D_i = \text{Demand for product} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$

In this exercise, you're willing to maximize profit given the cost of production.

```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
c = as.vector(data[5,2:3]) #This cost will be changed
max = as.matrix(data[4,2:3])
a = as.matrix(data[1:3,2:3])
p = as.matrix(data[1:3,4])

#Multiply all rows of j column of a[i,j] to c[i], in r for this to happen you have to change the vector to a square matrix
datatable(diag(c))
c = a %*% diag(c)
#It becomes:
datatable(c)

#Get back to model
model <- MIPModel() %>%
  # Variable of productW
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*(p[i] - sum_expr(c[i,j],j = 1:m)), i = 1:n), "max") %>%

  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j], j = 1:m)


#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

#Optimal Value
result$objective_value

#result solution
result$solution

```


## Week3 - IPs, MILPs & Network Design

### Lesson 1 - IPs & MILPs

#### Banner Chemicals ll


Question Info:assuming you can only produce in lots of 10 barrels.

```{r}

data <- tibble(Type = c("high","Supreme","Max"),
               Additive_1  = c(3,2,300), Additive_2 = c(1,3,280),
               Profit = c(80,200,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} \frac{x_{i} \cdot p_{i}}{Q} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \le \frac{P_i}{Q} \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le \frac{Max_j}{Q}, \ \ j \in M \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Product component in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Product requirement for stage j} \\
& P_i = \text{Capacity for Plant} \\
& Max_j = \text{Capacity for components} \\
& Q = \text{Lot} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$


```{r}
rm(x)

n = 2 #Number of products
m = 2 #Number of components
P = 110 #Plant Capacity

#subsetting data to R
max = as.matrix(data[3,2:3])
a = as.matrix(data[1:2,2:3])
p = as.matrix(data[1:2,4])

Model <- function(Q){

model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr((p[i] * x[i])/Q, i = 1:n), "max") %>%
  
  #Don't produce something you don't have demand for
  add_constraint(sum_expr(x[i], i = 1:n) <= P/Q) %>%
  
  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j]/Q, j = 1:m)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = F))

return(list(Objective = result$objective_value*Q^2,
            Variables = result$solution*Q))
}

#Question 1: Result when Lot = 10

Model(10)

#Question 2: Result when Lot = 5

Model(5)
```

#### GOnuts

```{r}

data <- tibble(Type = c("C_Ethiopia","C_Tanzania","C_Nigeria","Demand"),
               Ginko  = c(21,22.5,23,550), Kola = c(22.5,24.5,25.5,450),
               Capacity = c(425,400,750,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} \sum_{j=1}^{M} c_{i,j} \cdot x_{i,j} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \le C_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge D_i, \ \ j \in M \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Product component in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Product requirement for stage j} \\
& P_i = \text{Capacity for Plant} \\

& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$


```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
D = as.matrix(data[4,2:3])
a = as.matrix(data[1:3,2:3])
C = as.matrix(data[1:3,4])

#Modeling for Lesson and question, z is for the exercise, by changing the capacity of plant's constraint

Model <- function(z = NA){
  
  if (is.na(z)) {
   z = C
  } else {
   z = as.matrix(replicate(n = n, z))
  }
  
  #Get back to model
  model <- MIPModel() %>%
  # Variable of productW
  add_variable(x[i,j], i = 1:n, j = 1:m, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i,j]*a[i,j] ,j = 1:m, i = 1:n), "min") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i,j], i = 1:n) >= D[j], j = 1:m) %>%
  
  #Attend all demand
  add_constraint(sum_expr(x[i,j], j = 1:m) <= z[i], i = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = FALSE))
  
return(list(Obj_value = result$objective_value,
            Solution = result$solution))
  
}

#Question 1: Using the capacity of the lesson
Model()

#Question 2: All plants capacity are 350
Model(350)

```




#### QQ4 - Two Simple ILPs

##### Part l

```{r, warning = F}

data <- tibble(Type = c("Maximize", "C_1","C_2"),
               Var_1 = c(3,5,5), Var_2  = c(4,-3,10), RH = c(NA,8,63))

datatable(data)
```


```{r, warning= F}
rm(x)

n = 2 #Number of Variables & constraints

#subsetting data to R
a <- as.matrix(data[1,2:3])
c <- as.matrix(data[2:3,2:3])
RH <- as.matrix(data[2:3,4])

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*a[i], i = 1:n), "max") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*c[i,j], j = 1:n) <= RH[i], i = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```


##### Part 2

```{r, warning = F}

data <- tibble(Type = c("Maximize", "C_1","C_2","C_3"),
               Var_1 = c(2,-2,2,-2), Var_2  = c(3,1,0,2),
               State = c(NA,"<=","<=",">="), RH = c(NA,4,4,-2))

datatable(data)
```


```{r, warning= F}
rm(x)

n = 2 #Number of Variables & constraints


#subsetting data to R
a <- as.matrix(data[1,2:3])
c <- as.matrix(data[2:4,2:3])
RH <- as.matrix(data[2:4,5])


#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*a[i], i = 1:n), "max") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*c[j,i], i = 1:n) <= RH[j], j = 1:n) %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*c[j,i], i = 1:n) >= RH[j], j = 3)


#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```





#### GOnuts ll

```{r}

data <- tibble(Type = c("C_Ethiopia","C_Tanzania","C_Nigeria","Demand"),
               Ginko  = c(21,22.5,23,550), Kola = c(22.5,24.5,25.5,450),
               Capacity = c(425,400,750,NA), Plant_Cost = c(1500,2000,3000,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Minimize} & \ z = \sum_{i=1}^{N} \sum_{j=1}^{M} c_{i,j} \cdot x_{i,j} +
                          \sum_{j=1}^{M} f_j \cdot y_j \\
                          
\textbf{subject to} \\
    & \sum_{i=1}^{N} x_{i} \le C_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge D_i, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} -  f_j \cdot B_M \le 0\\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of Plant in N} \\
& j = \text{Demand in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Product requirement for stage j} \\
& P_j = \text{Capacity for Plant} \\
& B_M = \text{Big number}\\
& Max_j = \text{Capacity for components} \\
& x_{i} = \text{Quantity of product} \\

\end{align*}
$$


```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
D = as.matrix(data[4,2:3])
a = as.matrix(data[1:3,2:3])
C = as.matrix(data[1:3,4])
Pc = as.matrix(data[1:3,5])
B = sum(D) 

Model <- function(z = NA){
  
  if (is.na(z)) {
   z = C
  } else {
   z = as.matrix(replicate(n = n, z))
  }

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i,j], i = 1:n, j = 1:m, type = "integer", lb = 0) %>%
  
  #Binary Variable
  add_variable(f[i], i = 1:n, type = "binary") %>%

  # maximize profit
  set_objective(sum_expr(x[i,j]*a[i,j] ,j = 1:m, i = 1:n) +
                sum_expr(f[i]*Pc[i], i = 1:n), "min") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i,j], i = 1:n) >= D[j], j = 1:m) %>%
  
  #Attend all demand
  add_constraint(sum_expr(x[i,j], j = 1:m) <= z[i], i = 1:n) %>%
  
  #
  add_constraint(sum_expr(x[i,j], j = 1:m) - f[i]* B <= 0, i = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = FALSE))

return(list(Obj_value = result$objective_value,
            Solution = result$solution))

}

#Question 1: Using the capacity of the lesson
Model()

#Question 2: All plants capacity are 450
Model(450)

```




#### GOnuts lll


```{r}

data <- tibble(Type = c("C_Ethiopia","C_Tanzania","C_Nigeria","Demand"),
               Plant_Cost = c(1500,2000,3000,NA),
               Ginko  = c(21,22.5,23,550), Kola = c(22.5,24.5,25.5,450),
               Min_Cap = c(100,250,600,NA), Max_Cap = c(425,400,750,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Minimize} & \ z = \sum_{i=1}^{N} \sum_{j=1}^{M} c_{i,j} \cdot x_{i,j} +
                          \sum_{j=1}^{M} f_j \cdot y_j \\
                          
\textbf{subject to} \\
    & \sum_{i=1}^{N} x_{i,j} \le C_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} \ge L_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} \cdot a_{i,j} \ge D_i, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} -  f_j \cdot B_M \le 0\\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of Plant in N} \\
& j = \text{Demand in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Product requirement for stage j} \\
& P_j = \text{Capacity for Plant} \\
& B_M = \text{Big number}\\
& Max_j = \text{Capacity for components} \\
& x_{i} = \text{Quantity of product} \\

\end{align*}
$$


```{r, warning= F}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
D = as.matrix(data[4,3:4])
a = as.matrix(data[1:3,3:4])
L = as.matrix(data[1:3,5])
C = as.matrix(data[1:3,6])
Pc = as.matrix(data[1:3,2])
B = sum(D)

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i,j], i = 1:n, j = 1:m, type = "integer", lb = 0) %>%
  
  #Binary Variable
  add_variable(f[i], i = 1:n, type = "binary") %>%

  # maximize profit
  set_objective(sum_expr(x[i,j]*a[i,j] ,j = 1:m, i = 1:n) +
                sum_expr(f[i]*Pc[i], i = 1:n), "min") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i,j], i = 1:n) >= D[j], j = 1:m) %>%
  
  #Attend all demand
  add_constraint(sum_expr(x[i,j], j = 1:m) <= C[i], i = 1:n) %>%
  
  #
  add_constraint(sum_expr(x[i,j], j = 1:m) >= L[i]*f[i], i = 1:n) %>%
  
  #
  add_constraint(sum_expr(x[i,j], j = 1:m) - f[i]* B <= 0, i = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```


#### GOnuts lV


Operate with 2 Plants at maximum constraint.

$$
\begin{align*}

\textbf{Minimize} & \ z = \sum_{i=1}^{N} \sum_{j=1}^{M} c_{i,j} \cdot x_{i,j} +
                          \sum_{j=1}^{M} f_j \cdot y_j \\
                          
\textbf{subject to} \\
    & \sum_{i=1}^{N} x_{i,j} \le C_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} \ge L_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} \cdot a_{i,j} \ge D_i, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i,j} -  f_j \cdot B_M \le 0\\
    & \sum_{j=1}^{M} f_j \le K \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of Plant in N} \\
& j = \text{Demand in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Product requirement for stage j} \\
& P_j = \text{Capacity for Plant} \\
& B_M = \text{Big number}\\
& Max_j = \text{Capacity for components} \\
& x_{i} = \text{Quantity of product} \\
& K = \text{Maximum of Plants to be open} \\

\end{align*}
$$


```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components
K = 2

#subsetting data to R
D = as.matrix(data[4,3:4])
a = as.matrix(data[1:3,3:4])
L = as.matrix(data[1:3,5])
C = as.matrix(data[1:3,6])
Pc = as.matrix(data[1:3,2])
B = sum(D)

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i,j], i = 1:n, j = 1:m, type = "integer", lb = 0) %>%
  
  #Binary Variable
  add_variable(f[i], i = 1:n, type = "binary") %>%

  # maximize profit
  set_objective(sum_expr(x[i,j]*a[i,j] ,j = 1:m, i = 1:n) +
                sum_expr(f[i]*Pc[i], i = 1:n), "min") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i,j], i = 1:n) >= D[j], j = 1:m) %>%
  
  #Attend all demand
  add_constraint(sum_expr(x[i,j], j = 1:m) <= C[i], i = 1:n) %>%
  
  #TODO
  add_constraint(sum_expr(f[j], j = 1:m) <= K) %>%
  
  #TODO
  add_constraint(sum_expr(x[i,j], j = 1:m) >= L[i]*f[i], i = 1:n) %>%
  
  #TODO
  add_constraint(sum_expr(x[i,j], j = 1:m) - f[i]* B <= 0, i = 1:n)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```


### Lesson 2 - Networks & NLP

#### Recitations

##### Integer Linear Programming


```{r, warning = F}

data <- tibble(Type = c("Maximize", "C_1","C_2","C_3"),
               Var_1 = c(8,1,3,1), Var_2  = c(20,1,2,3), RH = c(NA,11,30,28))

datatable(data)
```


```{r, warning= F}
rm(x)

n = 2 #Number of Variables
m = 3 #Number of constraints

#subsetting data to R
a <- as.matrix(data[1,2:3])
c <- as.matrix(data[2:4,2:3])
RH <- as.matrix(data[2:4,4])

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*a[i], i = 1:n), "max") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*c[j,i], i = 1:n) <= RH[j], j = 1:m)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```


##### Capital Budgeting Problem


```{r, warning = F}

data <- tibble(Project = c("P1","P2","P3","P4","P5","Funds"),
               Exp_1 = c(5,4,3,7,8,25), Exp_2 = c(1,7,9,4,6,25),
               Exp_3 = c(8,10,2,1,10,25), Returns = c(20,40,20,15,30,NA))

datatable(data)
```


```{r, warning= F}
rm(x)

n = 5 #Number of Projects
m = 3 #Number of Years

#subsetting data to R
a <- as.matrix(data[1:5,5])
c <- as.matrix(data[1:5,2:4])
RH <- as.matrix(data[6,2:4])

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i], i = 1:n, type = "binary", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*a[i], i = 1:n), "max") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*c[i,j], i = 1:n) <= RH[j], j = 1:m)

#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```


##### Fixed Charge Problem

```{r, warning = F}

data <- tibble(Type = c("Shirts","Short","Pants","RH"),
               Labor = c(3,2,6,150), Cloth = c(4,3,4,160),
               Sales = c(12,8,15,NA), Fix_Cost = c(200,150,100,NA),
               Var_Cost = c(6,4,8,NA))

datatable(data)
```

```{r, warning= F}
rm(x)

n = 3 #Number of Products

#subsetting data to R
a <- as.matrix(data[1:3,2:3])
RH <- as.matrix(data[4,2:3])
p <- as.matrix(data[1:3,4])
c <- as.matrix(data[1:3,5:6])
M <- 1000

#Get back to model
model <- MIPModel() %>%
  # Variable of production
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%
  
  # Variable of fixed cost
  add_variable(y[i], i = 1:n, type = "binary", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(x[i]*(p[i]-c[i,2]) - y[i]*c[i,1], i = 1:n), "max") %>%

  #Attend all demand
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= RH[j], j = 1:n) %>%
  
  #Flow Constraint
  add_constraint(sum_expr(x[i], i = 1:n) <= y[i]*M, i = 1:n)
  

model$constraints


#Solve
result <- solve_model(model, with_ROI(solver = "glpk"))

#Optimal Value
result$objective_value

#result solution
result$solution

```