--- 
title: "A CTLSCx Exercises in R"
author: "Eduardo Almeida"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
github-repo: Eduardodudu/CTLSCx
url: 'https://github.com/Eduardodudu/CTLSCx'
description: "Trying to learn NERD Network design ?"
---

# CTLSC0x - Supply Chain Analytics

```{r include=FALSE}
library(ompr)
library(ROI)
library(ROI.plugin.glpk)
library(ompr.roi)
library(ggplot2)
library(plotly)
library(tibble)
library(DT)
library(tidyverse)
```


## Week2 - Unconstrained & Constrained Optimization

### Recitation

#### Unconstrained

Straigthforward exercises in modeling optimization without index vectors.

$$
\textbf{Equation 1:} \ \ Minimze \ \ y = x^2+2*x-3
$$

Since is an unconstrained quadratic optimization with a single variable, base R package optimizer can solve it

```{r}
f <- function(x){x^2+2*x-3}
result <- optimize(f,interval = c(-10,10), maximum = F)
print(result)

```

Let's check the result in the plot

```{r}
#Data created for plot, using mutate on the function
x <- seq(-10,10,1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$minimum)

```



$$
\textbf{Equation 2:} \ \ Maximize \ \ z =-x^2+2*x-y^2
$$

In this equation there is 2 variables, for this case it can be used the base R package optim.


```{r}
f <- function(x,y){-x[1]^2+2*x[1]-x[2]^2}
result <- optim(c(1, 1), f)
print(result)

```


#### Constrained

##### Banner Chemicals

```{r}

data <- tibble(Type = c("Profit","Plant","Additive A", "Additive B"), High = c(80,1,3,1),
                   Supreme = c(200,1,2,3), Capacity = c(NA,110, 300, 280))

datatable(data)

```



$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i\in M}^{ }p_i\cdot x_i \\
\textbf{subject to}\\
    & \sum_{i\in M} x_i \le\ C \\
    & \sum_{i\in M} a_{i,j} \cdot x_i \le\ A_j, \ \ \forall j \in N \\
    & x_i \ge 0 \\
    \\
    
\textbf{Where} \\

& i = \text{Products in M} \\
& j = \text{Additives in N} \\
& p_i = \text{Profit margin for product i} \\
& C = \text{Plant capacity} \\
& A_j = \text{Additive j available} \\
& a_{i,j} = \text{Quantity of additive j required per barrel of product i} \\
& x_i = \text{Quantity of product i to produce} \\

\end{align*} 
$$

```{r, warning=FALSE}
rm(x)

n = 2 #Number of products High and Supreme
C = filter(data, Type == "Plant") %>% select(Capacity) %>% .$Capacity
a = as.matrix(filter(data, Type %in% c("Additive A","Additive B")) %>% select(High, Supreme))
A = filter(data, Type %in% c("Additive A","Additive B")) %>% select(Capacity) %>% .$Capacity
p = filter(data, Type == "Profit") %>% select(High, Supreme) %>% gather() %>% .$value


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # minimize travel distance
  set_objective(sum_expr(p[i]*x[i], i = 1:n), "max") %>%

  # you cannot exceed the Plant Capacity
  add_constraint(sum_expr(x[i], i = 1:n) <= C) %>%
  
  #You cannot exceed aditivie capacity
  add_constraint(sum_expr(x[i] * a[j,i], i = 1:n) <= A[j], j = 1:n)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)

```


### Practice Problems

#### Handmade Baskets

$$
P = (x/73)^2-(x/95)+0.92
$$

```{r}
P <- function(x){(x/73)^2-(x/95)+0.92}

#Data created for plot, using mutate on the function
x <- seq(-100,200,1)
data <- data <- data.frame(x) %>%
        mutate(y = P(x))

min = data[data$y == min(data$y),] #cheating :P

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = min$y, x = min$x)

```




```{r}
#Part 1: Price Functio

#Question1: How much will it charge with 5 basket's order ?
P(5)

#Question2: What is the first derivative of the price function?
library(Ryacas)
x <- Sym("x")
s <- expression((x/73)^2-(x/95)+0.92)
deriv(s,x) #Didn't work TO DO

#Question3: Second Derivative
x <- Sym("x")
s <- expression(-1/95 + (2*x)/5329)
deriv(s,x) #worked

#Question4: Lowest Unit Price
result <- optimize(P,interval = c(-100,100), maximum = F)
print(result)

```

#### Maximizing Storage


$$
\begin{align*}
\textbf{Minimize} & \ Z = x*y \\
\textbf{st.} \\
& 2*x+2*y = 26.7
\end{align*}
$$


Simplifying $2*x+2*y = 26.7$ as $y = (26.7 - 2*x)/2$ is possible to change the optimization problem to a single variable:

$$
\begin{align*}
\textbf{Minimize} & \ Z\ =x\cdot\ \left(\frac{\left(26.7-2\cdot x\right)}{2}\right)\\
\end{align*}
$$


Solving:

```{r}
f <- function(x){x*((26.7 - 2*x)/2)}
result <- optimize(f,interval = c(-100,100), maximum = T)
print(result)

```


Let's check the result in the plot

```{r}
#Data created for plot, using mutate on the function
x <- seq(-100,100,0.1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)

```


#### Marketing MagicPuppy

$$
N = 1700 \left ( \frac{x}{90} - \left ( \frac{x}{90} \right ) ^2 \right )
$$



```{r}
f <- function(x){1700*(x/90-(x/90)^2)}

#Question1: First derivative
x <- Sym("x")
s <- expression(1700*(x/90-(x/90)^2))
deriv(s,x) #Didn't work TO DO


#Question2: How many units should be given away in the campaign in order to maximize its positive impact?
result <- optimize(f,interval = c(-100,200), maximum = T)
print(result)


x <- seq(-50,100,1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)

```

#### CureBase Cancer Institute


$$
W = \left ( m^{x^{2}} \right )\left ( n^{-x} \right )\left ( z^7 \right )
$$


```{r}
#Question what is the value of  x  at the maximum value of  W , given that  m=0.06 ,  n=0.14 ,  z=0.15 ?)
f <- function(x){0.06^(x^2)*(0.14^(-x))*0.15^7}
result <- optimize(f,interval = c(-1,1), maximum = T)
print(result)
```



```{r}
x <- seq(-1,1,0.1)
data <- data <- data.frame(x) %>%
        mutate(y = f(x))

#Ploting with a marker on the optimal solution
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>%
  add_markers(y = result$objective, x = result$maximum)
```

 
#### Santa’s Molding Company


```{r}

data <- tibble(Type = c("Profit","Time","Cubic Area", "index capacity"), Toy = c(25,13,33,75),
                   doll = c(25,15,25,70), Capacity = c(NA,1440, 3000, NA))

datatable(data)

```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i\cdot x_i \\
\textbf{subject to}\\
    & x_i \le\ c_i  \ \ \forall i \in N \\
    & \sum_{i=1}^{N} t_{i} \cdot x_i \le\ 1440 \\
    & \sum_{i=1}^{N} a_{i} \cdot x_i \le\ 3000 \\
    & x_i \ge 0 \\
    \\
    
\textbf{Where} \\

& i = \text{Products in N} \\
& p_i = \text{Profit margin for product i} \\
& c_i = \text{Product daily capacity} \\
& t_i = \text{Time to mold product i} \\
& a_i = \text{Cubic inches of product i} \\
& x_i = \text{Quantity of product i to mold} \\

\end{align*} 
$$

If we look closer on the dataframe, we can set constraints 2 and 3 as one with a matrix including capacity constraints, but for math notation I've set them differently since they are 2 different types of entities.


```{r}
rm(x)

n = 2 #Number of products Toy and doll
c = filter(data, Type == "index capacity") %>% select(Toy,doll) %>% gather() %>% .$value #index_capacity
a = as.matrix(filter(data, Type %in% c("Time","Cubic Area")) %>% select(Toy,doll)) #constraints left hand
C = filter(data, Type %in% c("Time","Cubic Area")) %>% select(Capacity) %>% .$Capacity #constraints right hand
p = filter(data, Type == "Profit") %>% select(Toy,doll) %>% gather() %>% .$value #profit



model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # minimize travel distance
  set_objective(sum_expr(p[i]*x[i], i = 1:n), "max") %>%

  # you cannot exceed the Plant Capacity
  add_constraint(x[i] <= c[i], i= 1:n) %>%

  #You cannot exceed aditivie capacity
  add_constraint(sum_expr(x[i] * a[j,i], i = 1:n) <= C[j], j = 1:n)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)

```


Let's solve the questions

```{r}
P <- function(x,y){25*x+25*y}
#Question 1: Graphical selection
#Question 2: What is the value of x = 70 and y= 30
P(70,30)

#Question 3: What is the value of x = 50 and y= 53
P(50,53)

#Question 3: What is the value of x = 50 and y= 53
P(20,75)

```

#### Crazy Cereal

```{r}

data <- tibble(Type = c("Sugary","Regular","Capacity"), Sugar = c(0.66,0.21,2000),
                   Corn_Flake = c(0.34,0.79,4000), Profit = c(0.94,0.82,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le C_j, \ j \in M \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    \\
    
\textbf{Where} \\

& i = \text{Products in N} \\
& j = \text{Components of N = M} \\
& p_i = \text{Profit margin for product i} \\
& a_{i,j} = \text{components for product i} \\
& C_j = \text{Component capacity} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake

#Let's do some roots R instead of dplyr :)
C = as.matrix(data[data$Type == "Capacity",c("Sugar", "Corn_Flake")])
a = as.matrix(data[data$Type %in% c("Sugary","Regular"),c("Sugar", "Corn_Flake")])
p = as.matrix(data$Profit[1:2])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(p[i] * x[i], i = 1:n), "max") %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:m) <= C[j], j = 1:n)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)


```

#### Jim's Meat Packing Company

```{r}

data <- tibble(Cut = c("Chuck","Sirloin","Restriction"), Lean_Meat = c(0.09,0.6,0.3),
                   Fat_Meat = c(0.02,0.06,0.05), Cost = c(9.3,8.4,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=1 \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=2 \\
    & \sum_{i=1}^{N} x_{i} \ge 50 \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
    \\
    
\textbf{Where} \\

& i = \text{Type of cut in N} \\
& j = \text{Type of meat} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& R_j = \text{Restriction capacity} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake

#Another way of subsetting data in R
R = as.matrix(data[3,2:3])
a = as.matrix(data[1:2,2:3])
c = as.matrix(data[1:2,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #yay
  add_constraint(sum_expr(x[i], i = 1:n) >= 50) %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

  
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)


```


I couldn't find a proper function on shadow price calculation on ompr package, so lets try to calculate it ourselves!

```{r}

ModelIter <- function(z){
  
  model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #yay
  add_constraint(sum_expr(x[i], i = 1:n) >= z) %>%

  #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2) %>%
    
  solve_model(with_ROI(solver = "glpk", verbose = F))
  
  return(model$objective_value)
  
}

Iter_Results <- tibble(constraint = seq(48,55,1)) %>%
                mutate(Obj_value = as.numeric(map(constraint, ModelIter)),
                       Marginal_value = round((Obj_value-lag(Obj_value,1)/(constraint-lag(constraint,1))),2))
datatable(Iter_Results)

mean(Iter_Results$Marginal_value[2:length(Iter_Results)])

```

Marginal Value set as 8.85 got us an error, the answer was 8.63, would be good to understand what was made wrong in here.


#### John's Shipping Company

```{r}

data <- tibble(Fuel_Type = c("O_Type","D_Type","Restriction"), Hydrogen_Conc = c(45,90,60),
                   Oxygen_Conc = c(15,6,9), Cost = c(1.05,1.34,NA))
datatable(data)

```


$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=1 \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le \sum_{i=1}^{N} R_j \cdot x_{i}, \ \ j \in N=2 \\
    & \sum_{i=1}^{N} x_{i} \ge G \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
    \\
    
\textbf{Where} \\

& i = \text{Type of cut in N} \\
& j = \text{Type of meat} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& R_j = \text{Restriction capacity} \\
& x_{i} = \text{Quantity of product} \\
& G = \text{Minimum to be produced} \\

\end{align*} 
$$



```{r}
rm(x)

n = 2 #Number of products Sugary and Regular
m = 2 #Number of components Sugar and Corn_Flake
z = 10000 #minimum to be produced

#Another way of subsetting data in R
R = as.matrix(data[3,2:3])
a = as.matrix(data[1:2,2:3])
c = as.matrix(data[1:2,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #Produce at least the minimum required
  add_constraint(sum_expr(x[i], i = 1:n) >= z) %>%

  #Minimum component required
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) >= sum_expr(x[i], i = 1:n)*R[j], j = 1) %>%

   #You cannot exceed component capacity
  add_constraint(sum_expr(x[i] * a[i,j], i = 1:n) <= sum_expr(x[i], i = 1:n)*R[j], j = 2)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)


```


#### A New Office Plant

```{r}

data <- tibble(Type = c("Product_1","Product_2","Product_3", "Min", "Max"), Nitrogen  = c(360,380,310,1800,2200),
                   Potassium = c(30,20,20,100,100), cost = c(1.59,2.19,2.99,NA,NA))
datatable(data)

```



$$
\begin{align*}

\textbf{Minimize} & \ y = \sum_{i=1}^{N} c_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \ge Min_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le Max_j, \ \ j \in M \\
    & x_{i} \ge 0 \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Type of component in M} \\
& c_i = \text{Cost for product i} \\
& a_{i,j} = \text{components for product i} \\
& Min_j = \text{Restriction of component} \\
& Max_j = \text{Restriction of component} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$



```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
min = as.matrix(data[4,2:3])
max = as.matrix(data[5,2:3])
a = as.matrix(data[1:3,2:3])
c = as.matrix(data[1:3,4])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "continuous", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(c[i] * x[i], i = 1:n), "min") %>%
  
  #Produce at least the minimum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) >= min[j], j = 1:m) %>%
  
  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j], j = 1:m)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

#Duals
get_column_duals(result)


#Question: How many mg of Nitrogen would your solution provide to the plant every day? (Please round your answer to integer number)
result$solution[1]*a[1,1]+result$solution[2]*a[2,1]

```


#### Rochak's Ink Company

```{r}

data <- tibble(Type = c("P_Cyan","P_Magenta","P_Yellow","Max"),
               Stage1  = c(0.44,0.51,0.5,17*60^2), Stage2 = c(0.55,0.6,0.59,23*60^2),
               Demand = c(84000,72000,93000,NA), Profit = c(0.6,0.72,0.62,NA))

datatable(data)
```


$$
\begin{align*}

\textbf{Maximize} & \ y = \sum_{i=1}^{N} p_i \cdot x_{i} \\
\textbf{subject to}\\
    & \sum_{i=1}^{N} x_{i} \cdot a_{i,j} \le Max_j, \ \ j \in M \\
    & \sum_{i=1}^{N} x_{i} \le D_i \\
    & x_{i} \ge 0, \ x \in \mathbb{Z} \\
    
\textbf{Where} \\

& i = \text{Type of product in N} \\
& j = \text{Production stage in M} \\
& p_i = \text{Profit for product i} \\
& a_{i,j} = \text{Time requirement for product i in stage j} \\
& Max_j = \text{Restriction of production time} \\
& D_i = \text{Demand for product} \\
& x_{i} = \text{Quantity of product} \\

\end{align*} 
$$

In this exercise, you're willing to maximize profit given the demand. You don't have to fulfill all of the demand, that's why the $ \sum_{i=1}^{N} x_{i} \le D_i$ constraint.



```{r}
rm(x)

n = 3 #Number of products
m = 2 #Number of components

#subsetting data to R
D = as.matrix(data[1:3,4])
max = as.matrix(data[4,2:3])
a = as.matrix(data[1:3,2:3])
p = as.matrix(data[1:3,5])


model <- MIPModel() %>%
  # Variable of profit
  add_variable(x[i], i = 1:n, type = "integer", lb = 0) %>%

  # maximize profit
  set_objective(sum_expr(p[i] * x[i], i = 1:n), "max") %>%
  
  #Don't produce something you don't have demand for
  add_constraint(x[i] <= D[i], i = 1:n) %>%
  
  #Dont exceed maximum required
  add_constraint(sum_expr(x[i]*a[i,j], i = 1:n) <= max[j], j = 1:m)


model$constraints

#Solve
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

#Optimal Value
result$objective_value

#result solution
result$solution

```

#### Pasteur Cheese Factory



